---
vignette: |
  %\VignetteIndexEntry{ArMag} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
author: "Dufresne Philippe"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
  vignette: >
   %\VignetteIndexEntry{ArMag}
   %\VignetteEngine{knitr::rmarkdown}
   %\VignetteEncoding{UTF-8}
    
title: "Calcul statistique sur les composantes magnétiques avec le package ArMag"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width=7, fig.height = 5, fig.align = 'center',
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This package is an extraction of the functions of the calculation software on magnetic components developed at the Rennes Archaeomagnetism Laboratory, hosted at the Geosciences-Rennes laboratory.

## Installing ArMag

ArMag will run in Windows, Mac OS X or Linux. To install ArMag you first need to install [R](https://www.r-project.org). I would also recommend installing [Rstudio](http://www.rstudio.com) as a nice desktop environment for using R. Once in R you can type:

```{r,eval=FALSE}
# install.packages('ArMag')
```


at the R command prompt to install ArMag. If you then type:

```{r}
# library(ArMag)
```

## Installing ArMag via GitHub
```{r echo = TRUE}

if (!require(devtools))
 {install.packages("devtools")}

devtools::install_github("chrono35/ArMag", force = TRUE)
library("ArMag") #, lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
```

it will load in all the ArMag functions.

## Loading Rennes' AM file
You need two functions, the first one reads the information, corresponding to the headers of each sample, the second one reads the measurements of each sample.
```{r}

file.AM <- "../examples/14039C.AMP"

mes <- NULL
mes.info <- read.AM.info (file.AM)
mes <- read.AM.mesures(file.AM)

```

## Empty file generation
Function used before making the measurements.
Warning, this function overwrites the existing file
```{r genere AMD}
file.AM <- "../examples/test.txt"
list.ech <- c("1T",   "17T",  "35T",  "47T",  "50T",  "59T",  "83T",  "87T",  "89T",  "94T",  "100T", "102T", "103T" )
genere.AMD(file.AM, list.ech)
```

## Plot of NRN values
A simple, R command allows to select the steps with the value 0, corresponding to step 0N0
```{r lambert}
select <- NULL
# select$I <- mes$I[mes$step.value== 0]
# select$D <- mes$I[mes$step.value== 0]
select <- mes[mes$step.value== 0,]
lambert(select, inc.lim = c(0,90))
```

## mcFadden statistic on NRN value
Example of the calculation of the mcFadden statistic on the selected samples
```{r stat mcFadden}
stat.mcFadden(select)

```

## Extraction of measurements corresponding to a sample using its name
```{r selection}
mes.ech <- NULL
mes.ech <- extract.mesures.specimen.name("10P1", mes)
```

## Zijderveld plot
```{r fonction plot.zjiderveld1, fig.show='hold'}
par(pty="s") # force une figure carré
zijderveld1(mes.ech$X, mes.ech$Y, mes.ech$Z) 
par(mfrow = c(1,2), pty="m", cex.lab = 0.5, cex.axis = 0.6) # separated into 2 columns and restored a maximum size figure
# cex.lab set the text size of the steps
zijderveld1(mes.ech, legend.pos = "topright")
zijderveld2(mes.ech, pt.names = NULL)
```

## Plot the schmitt-Lambert projection
Each function using different parameters
```{r comparatif fonction plot, fig.show='hold'}
lambert.XYZ.specimen(mes.ech)
# remove anisotropîe step
mes.ech.ssAni<-remove.step(mes.ech)
lambert.ID.specimen(mes.ech.ssAni)
```

## demagnetizing and partial component

```{r, fig.show='hold'}
mes.sel2 <- extract.mesures.specimen.name("1P1", mes)
mes.sel3 <- extract.mesures.specimen.name("16P1", mes)
demag(mes.sel2, step.J0 = NULL)
demag(mes, step.J0 = 0, pt.col = rainbow(length(mes.info$name)))
demag(rbind(mes.sel2, mes.sel3), normalize = TRUE)

partial.component(mes.sel2$X, mes.sel2$Y, mes.sel2$Z)
```

## Synthetic view of a sample
The zijderveld1.T1T2 and zijderveld2.T1T2 functions remove the anisotropy steps by default

```{r assemblé, fig.show='hold', fig.cap="Synthetic view"}

par(mfrow = c(2, 2), cex.lab = 0.7, cex.axis = .7, cex = 0.7, cex.main = 1, cex.sub = 0.1,
    mai = c(0.5, 0.5, 0.7, 0.3), oma = c(0, 1, 1, 1))#, pty ="s" )
zijderveld1.T1T2(mes.sel2)
zijderveld2.T1T2(mes.sel2)

# removal of anisotropy steps
mes.sel2 <- remove.step(mes.sel2, verbose = FALSE) 
lambert(mes.sel2, inc.lim = c(0, 90))
demag(mes.sel2, step.J0 = NULL)
```

## Arai - intensity plot
```{r}
file.INT <- "../examples/INT_example.AMD"

mesINT <- NULL
mesINT <- read.AM.mesures(file.INT)
mesINT.info <- read.AM.info (file.INT)
mes.sel<- extract.mesures.specimen.name("40001B_11B1", mesINT)
# reference 
# Coe 1978 : DOI: 10.1029/JB083iB04p01740
# Prévost et Al. 1985 DOI: 10.1029/JB090iB12p10417
par(pty="s", "xaxp")
relative = FALSE
verbose = TRUE
show.plot = TRUE
vol=10.8
TH = 60
aim.coef = 1E-10*1E6/vol #1E6
show.step.value = FALSE
R.mark = 'R'  # Positive pTRM
V.mark = 'V'  # Negative pTRM
P.mark = 'P'  # pTRM check
L.mark = "L"  # sLow cooling 
Q.mark = "Q"  # Quick cooling
pt.col = "blue"
loop.col = "forestgreen"
step.J0 = "20N0" # ou NULL

begin.step.value = 0
end.step.value = 700

par(pty="s", "xaxp")
arai(mes.sel, begin.step.value = 250, end.step.value = 700, aim.coef = 1E-10*1E6/vol)
```

# Sun azimuth
```{r sun.azimuth}
sun.azimuth (10, 11, 2019, 11, 19, seconde=0, 48, longmin=0, longsec=0, 45, latmin=0, latsec=0)
```

# igrf13syn
  This is a synthesis routine for the 13th generation [IGRF](https://www.ngdc.noaa.gov/IAGA/vmod/igrf.html) as agreed in December 2019 by IAGA Working Group V-MOD. It is valid 1900.0 to 2025.0 inclusive. Values for dates from 1945.0 to 2015.0 inclusive are definitive, otherwise they are non-definitive.
  Reference:

 [Thébault, E., Finlay, C.C., Beggan, C.D. et al. International Geomagnetic Reference Field: the 12th generation. Earth Planet Sp 67, 79 (2015).  https://doi.org/10.1186/s40623-015-0228-9](https://earth-planets-space.springeropen.com/articles/10.1186/s40623-015-0228-9)

  
```{r}
isv <- 1
date <- 2000
itype <- 1
alt <- 50
colat <- 90 - 45
elong <- 10

igrf13syn(isv=isv, date=date, itype=itype , alt=alt, colat= colat, elong=elong)
```
